---
title: "Discrimination"
author: "David Kane"
format: html
execute: 
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
```

## From Monday's Class

### Background Information

Bertrand, Marianne and Sendhil Mullainathan. 2004. "Are Emily and Greg More Employable Than Lakisha and Jamal? A Field Experiment on Labor Market Discrimination." American Economic Review, 94 (4): 991–1013. Data cleaned and discussed in "Quantitative Social Science: An Introduction" by Kosuke Imai.

"We study race in the labor market by sending fictitious resumes to help-wanted ads in Boston and Chicago newspapers. To manipulate perceived race, resumes are randomly assigned African-American- or White-sounding names. White names receive 50 percent more callbacks for interviews. Callbacks are also more respon- sive to resume quality for White names than for African-American ones. The racial gap is uniform across occupation, industry, and employer size. We also find little evidence that employers are inferring social class from the names. Differential treatment by race still appears to still be prominent in the U.S. labor market."

### Discussion

Subject: Discrimination

Broad question: Race and employment in the United States

Specific question: Effect of racially-coded names on call-backs

Preceptor Table: 
  
  * Units: individuals (or resumes ?) 
  
  * (Potential) Outcomes: call-back if black name, call-back if white name   
  
  * Causal/predictive: a causal model  
  
  * Covariates: sex, zip code (not shown in table)  
  
  * Treatment: black name on resume versus white name on resume 

## Example Preceptor Table
  
```{r}
tibble(ID = c("1", "2", "...", "10", "11", "...", "N"),
       attitude_after_control = c("Yes*", "No", "...", "No*", "Yes*", "...", "Yes"),
       attitude_after_treated = c("Yes", "Yes*", "...", "No", "Yes", "...", "No*"),
       treatment = c("Black", "White", "...", "Black", "Black", "...", "White")) |>
  
  gt() |>
  tab_header(title = "Preceptor Table") |> 
  cols_label(ID = md("ID"),
             attitude_after_control = md("Callback if White Name"),
             attitude_after_treated = md("Callback if Black Name"),
             treatment = md("Treatment")) |>
  tab_style(cell_borders(sides = "right"),
            location = cells_body(columns = c(ID))) |>
  tab_style(style = cell_text(align = "left", v_align = "middle", size = "large"), 
            locations = cells_column_labels(columns = c(ID))) |>
  cols_align(align = "center", columns = everything()) |>
  cols_align(align = "left", columns = c(ID)) |>
  fmt_markdown(columns = everything()) |>
  tab_spanner(label = "Potential Outcomes", columns = c(attitude_after_control, attitude_after_treated)) |>
  tab_spanner(label = "Covariate", columns = c(treatment))
```

## Questions for Second Class

<!-- There are no "right" answers. Each person/group might have a different goal . . . -->

> What are the units, precisely?

Resumes.
All the people sending out resumes for entry-level positions in the US during 2020 -- 2030 who have names which are associated with a race. So, N is the hundreds of thousands (only people with specific names of races).

> What is the moment in time, precisely?

2020's

> What is one reason why validity might not hold?

- Biased sampling through targeting companies by sending specific resumes to each (biased data)
- Real World Events (Black Lives Matter)
- are still the names from 2004 black or white-coded?
- resumes now look completely different to 2001

<!-- Assume validity holds, and we can create the Population Table -->

> Describe the Population Table in words

The Population data is the data we want to have, the data which we actually have, and all the other data from that same population. It is combined with the Preceptor Table, data, other population data, and sources.

every row: unit + time

> What is one reason why stability might not hold?

A real-world event or governmental regime in the US may have impacted the US since 2001 to 2024. There are many examples. Such can be the BLM, Democratic or Republican President, etc.

> What is one reason why representativeness might not hold?

No specific companies - lesser known companies
Specific names?
Boston cannot fully represent the US

> What is one reason why unconfoundedness might not hold?

The ambiguity of the treatment and the randomness of the treatment values.

Resumes were sent to companies cannot be completely random. 

> Describe in words what the Preceptor Table would look like if we were trying to, instead, predict whether or not someone would get a callback?

We would not have a treatment column, and the data would be a lot more oriented towards being predictive towards the Nth ID.

<!-- If time, load up and explore the data.  -->
  